The Geometry of Archaeological Thought: Analysing discplinary change using word embeddings from conference abstracts

### 2020-05-14 documents/paper.Rmd
work on method section

- something about basic steps
-- transform text into dfm, define dfm...

- word embeddings (cite our readings)
-- a few sentences to introduce the reader to how WE have been used in other recent studies. this is a little bit background. '...recently used in many social science fields to...' tell the reader what was the finding of some key studies using WE that have inspired us.

@@

-- then some more detail about the mechanics of the method, as we are using it. Say how what we did works in general.

In this article, we trained word embeddings with the GloVe model using th text2vec package [@Selivanov2016]. The GloVe model calculates ... creates the co-occurrence matrix.

- temporal referencing plots (cite our readings)

>>>>

- Basic steps for matrix construction

We transform text into a document-feature matrix (dfm), where rows represent PDFs and columns represent features, the terms extracted from PDFs, and values are frequencies of certain features for each PDFs. We selected the features that occur at least three times for later construction of a word co-occurrence matrix. In this article, we trained word embeddings with the GloVe model using th text2vec package [@Selivanov2016]. The GloVe model calculates how often one word appear with the presence of another word to construct a co-occurrence matrix. This method involves setting an area defined by window in which we can detect the context terms with the target term in the middle position.

@@

- temporal referencing plots (cite our readings)
The word embeddings method is applied to explore temporal changes of vectors to present a historical trend. For example, @Garg2018 explore the historical trends in stereotypes and attitudes toward women and ethnic minorities in the 20th and 21st centuries in the US. They found both gender and ethnic occupation biases significantly correlated with the actual occupation participation rates. Those changes by negative words in the descriptions of genders and ethnic groups over time are associated with the women’s movement in the 1960s–1970s and Asian-American population growth in the 1960s and 1980s. Another study , @Kozlowski analyze the cultural dimensions of social class and how they evolve over the twentieth century in the US. They found the relationships between the cultural dimensions of class remain stable semantic structure over the century, but the individual words, such as employment, status, education, cultivation, on those dimensions shift over the decades. They also found affluence and status serve as cultural mediators between two clusters, education, cultivation, and morality on one side, and employment and ownership on the other side.

```{r temporal referencing plots}
# Words per year

library(tidyverse)

# starting with https://quanteda.io/articles/quickstart.html
library("quanteda")

# get all txt files of abstracts
abstract_txt_files <- 
  fs::dir_ls(here::here("data/derived-data/"), 
             recurse = TRUE, 
             regexp = '\\.txt$')

all_txts <- tolower(readtext::readtext(abstract_txt_files))
names(all_txts) <- map_chr(names(all_txts), ~str_match_all(.x, "\\d{4}")[[1]][1,1])

# word count checking
all_txts_word_count_per_year <- 
map_int(all_txts, 
       stringi::stri_count_words) %>% 
  stack %>% 
  as_tibble() %>% 
  rename(words = values,
         year = ind) %>% 
  mutate(year = as.numeric(as.character(year)))

# count all words for each year
all_txts_c <- corpus(all_txts)
all_txts_c_summary <- 
  summary(all_txts_c) %>%   
  mutate(year = parse_number(Text))

# make a dfm
all_txts_c_dtm <- 
  dfm(all_txts_c,
      remove_punct = TRUE)

# plot of all words in each year
plot_all_words <- 
  ggplot(data = all_txts_c_summary, 
         aes(x = year , 
             y = Tokens / 1000, 
             group = 1)) +
  geom_col() +
  scale_x_continuous(labels = c(seq(1940, 2020, 2)), 
                     breaks = seq(1940, 2020, 2),
                     name = "Year") +
  scale_y_continuous(name = "Total word count per year (x 1,000)",
                     labels = scales::comma) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5))

# Our key words over time

keywords <- 
  c("mechanism", 
    "theory", 
    "hypothesis", 
    "model", 
    "inference")

dfm_keywords <- 
  dfm_select(all_txts_c_dtm, 
             pattern = keywords, 
             selection = "keep")

dfm_keywords_tbl <- 
  convert(dfm_keywords, to = "data.frame") %>% 
  pivot_longer(-document, 
               names_to = "keyword",
               values_to = "n") %>% 
  mutate(year = parse_number(document))

# compute proportion of all words per year
dfm_keywords_tbl_prop <- 
dfm_keywords_tbl %>% 
  left_join(all_txts_c_summary) %>% 
  mutate(prop = n / Tokens ) %>% 
  group_by(keyword) %>% 
  mutate(sum_the_word = sum(n)) %>% 
  mutate(keyword_n = str_c(keyword, " (n = ", sum_the_word, ")"))

# plot of keywords as a proportion of all words per year
ggplot(data = dfm_keywords_tbl_prop, 
       aes(x = year , 
           y = prop)) +
  geom_point(alpha = 0.3) +
  geom_smooth(span = 0.475,
              se = FALSE) +
  facet_wrap( ~ keyword_n, 
              ncol = 1,
             scales = "free_y") +
  scale_x_continuous(labels = c(seq(1960, 2020, 2)), 
                     breaks = seq(1960, 2020, 2),
                     name = "Year") +
  scale_y_continuous(name = "Proportion of all words in all abstracts per year",
                     labels = scales::comma) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5))

```
